---
title: "MA 331 HW 1"
author: "Jack Byrne"
date: "2023-09-07"
output:
  html_document:
    df_print: paged
---

#HW #1 Due 11:59 PM Monday September 11, 2023

All parts of this homework should be submitted to Canvas. Submit the answers to Question 1 and Question 2 parts A and B using handwritten scanned notes or as a latex/typed up pdf file. Submit the answer to Question 2 Part C and Question 3 as the pdf file generated when knitting this report. If need be, combine these pdfs into a single file

##Question 1
Let X be a random variable which takes on the following values:

-1 with probability 1/3
0 with probability 1/3
1 with probability 1/3

Let Y be a random variable that is independent of X but that has the same probability distribution as X

Let Z be a random variable (also independent of X and Y) that takes on the values double that of X, that is, it has the pmf:

-2 with probability 1/3
0 with probability 1/3
2 with probability 1/3


###A) What is the Distribution of X + Y? (Hint: because X and Y are independent, the probability that we draw X to be 0 and Y to be 1 is equal to the product of their individual probabilities, that is, 1/3 * 1/3 = 1/9. In this case, their sum would be 1. However, is that the only way to get a sum of 1?)

 xy -1   0    1
-1 1/9  1/9  1/9

 0 1/9  1/9  1/9
 
 1 1/9  1/9  1/9
 
Domain of X+Y = [-2,-1,0,1,2]
P(X+Y = -2) = 1/9 *1 = 1/9
P(X+Y = -1) = 1/9 *2 = 2/9
P(X+Y =  0) = 1/9 *3 = 1/3
P(X+Y =  1) = 1/9 *2 = 2/9
P(X+Y =  2) = 1/9 *1 = 1/9

###B) What is the Distribution of X + Z?

 xz -2   0    2
-1 1/9  1/9  1/9

 0 1/9  1/9  1/9
 
 1 1/9  1/9  1/9
 
Domain of X+Y = [-3, -2,-1,0,1,2, 3]
P(X+Y = -3) = 1/9 *1 = 1/9
P(X+Y = -2) = 1/9 *1 = 1/9
P(X+Y = -1) = 1/9 *2 = 2/9
P(X+Y =  0) = 1/9 *1 = 1/9
P(X+Y =  1) = 1/9 *2 = 2/9
P(X+Y =  2) = 1/9 *1 = 1/9
P(X+Y =  3) = 1/9 *1 = 1/9

###C) Find the expected values of X, Y, Z, and X + Y. Show your work

E(X) = (1/3)*-1 + (1/3)*0 + (1/3)*1
     = -1/3 + 0 + 1/3
     = 0
     
E(Y) = (1/3)*-1 + (1/3)*0 + (1/3)*1
     = -1/3 + 0 + 1/3
     = 0

E(Z) = (1/3)*-2 + (1/3)*0 + (1/3)*2
     = -2/3 + 0 + 2/3
     = 0   
     
E(X+Y) = (1/9)*-2 + (2/9)*-1 + (1/3)*0 + (2/9)*1 + (1/9)*2
       = -2/9 + -2/9 + 0 + 2/9 + 2/9
       = 0
  
###D) Find the variances of X, Y, Z, and X + Y. Show your work

Var(X) = E(X^2) - [E(X)]^2

E(X^2) = (1/3)*-1^2 + (1/3)*0^2 + (1/3)*1^2
       = 1/3 + 0 + 1/3
       = 2/3

Var(X) = E(X^2) - [E(X)]^2 
       = 2/3 - 0^2 
       = 2/3
     
E(Y^2) = (1/3)*-1^2 + (1/3)*0^2 + (1/3)*1^2
       = 1/3 + 0 + 1/3
       = 2/3
       
Var(Y) = E(Y^2) - [E(Y)]^2 
       = 2/3 - 0^2 
       = 2/3

E(Z^2) = (1/3)*-2^2 + (1/3)*0^2 + (1/3)*2^2
       = 4/3 + 0 + 4/3
       = 8/3
     
Var(Z) = E(Z^2) - [E(Z)]^2 
       = 8/3 - 0^2 
       = 8/3

Var(X+Y) = Var(x) + Var(Y)
         = 2/3 + 2/3
         = 4/3

##Question 2
###A) Find the expected value of a Continuous Random Variable that has the following distribution:

f(x) = 1/2     for 0 <= x <= 2  
        0      elsewhere
        
E(X) = ∫−∞ ∞ x(f(x))dx

E(x) = ∫0 2 x(1/2)dx
     = 1/2 (∫0 2 x dx)
     = 1/2[x^2/2]0 2
     = 1/2 [4/2 - 0/2]
     = 1/2 * 2
     = 1
        
###B) Find the variance of the same distribution

Var(X) = E(X^2) - [E(X)]^2

E(X^2) = ∫−∞ ∞ x^2(f(x))dx

E(X^2) = ∫0 2 x^2(1/2)dx
       = 1/2 ∫0 2 x^2 dx
       = 1/2 [x^3 /3]0 2
       = 1/2 [8/3 - 0/3]
       = 1/2 * 8/3
       = 4/3
       
Var(X) = E(X^2) - [E(X)]^2
       = 4/3 - 1^2
       = 4/3 - 1
       = 1/3

###C) Sample from this distribution 1000 times, and give the empirical mean and variance. You may use the function runif(1000, 0, 2)
        
```{r}
samples <- runif(1000, 0, 2)

emp_mean <- mean(samples)

variance <- var(samples)

print(c("Empirical mean:", emp_mean))
print(c("Variance:", variance))

```


###A) Sample 10,000 samples from an exponential distribution with rate 1, and store the sample in a variable called expData and plot the histogram of these sampled values. You can use the function rexp(). Describe how the histogram looks to you

```{r}
expData <- rexp(10000, rate = 1)

hist(expData)
```
Description:
The histogram is skewed to the right and illustrates an exponential decay. The data is clustered at 0. The values range from 0 to 5.

###B) For each of 500 replicates, Sample 50 samples from an exponential distribution with rate 1, and store the mean of these samples for each replicate in a vector called expSample50. Plot the histogram of expSample50, and describe how it looks, especially in comparison with the histogram from your sample of size 10,000 from part A

```{r}
expSample50 <- rep(0, 500)

for (i in 1:length(expSample50)) {
  sample_50 <- rexp(50, rate = 1)
  expSample50[i] <- mean(sample_50)
}

hist(expSample50)

```
Description:
This histogram is more symmetrical than the one in part A, but is still slightly skewed right. The shape resembles a normal distribution. The data is clustered around 1.0-1.1 and ranges from 0.7-1.4.

###C) For each of 500 replicates, Sample 1000 samples from an exponential distribution with rate 1, and store the mean of these samples for each replicate in a vector called expSample1000. Plot the histogram of expSample100, and describe how it looks, especially in comparison with the histograms from parts A and B. 

```{r}
expSample1000 <- rep(0, 500)

for (i in 1:length(expSample1000)){
  sample_1000 <- rexp(1000, rate = 1)
  expSample1000[i] <- mean(sample_1000)
}

hist(expSample1000)

```
Description:
The histogram is the most symmetrical so far, resembling a normal distribution. The data is clustered around 1, and ranges from 0.9-1.1

###D) Compare the means and standard deviations of expData, expSample50, and expSample1000, and add any comments you may have
```{r}
expData <- rexp(10000, rate = 1)

expSample50 <- rep(0, 500)

for (i in 1:length(expSample50)) {
  sample_50 <- rexp(50, rate = 1)
  expSample50[i] <- mean(sample_50)
}


expSample1000 <- rep(0, 500)

for (i in 1:length(expSample1000)){
  sample_1000 <- rexp(1000, rate = 1)
  expSample1000[i] <- mean(sample_1000)
}

mean_expData <- mean(expData)
sd_expData <- sd(expData)

mean_expSample50 <- mean(expSample50)
sd_expSample50 <- sd(expSample50)

mean_expSample1000 <- mean(expSample1000)
sd_expSample1000 <- sd(expSample1000)

print(c("Mean of expData: ", mean_expData))
print(c("Mean of expSample50: ", mean_expSample50))
print(c("Mean of expSample1000: ", mean_expSample1000))

print(c("Standard Deviation of expData: ", sd_expData))
print(c("Standard Deviation of expSample50: ", sd_expSample50))
print(c("Standard Deviation of expSample1000: ", sd_expSample1000))

```
Description:
This is what I calculated:
Mean of expData: 1.00441800462773
Mean of expSample50: 0.994658051771891    
Mean of expSample1000: 0.99800005492377
Standard Deviation of expData: 1.02849423231158"               
Standard Deviation of expSample50:  0.139759162761012           
Standard Deviation of expSample1000: 0.0313872667872841

The three means remained relatively the same around 1. However the standard deviations decreased with sample size.

