---
title: "HW_2"
output: pdf_document
date: "2023-09-18"
---

Jack Byrne

#HW #2 Due 11:59 PM Monday September 18, 2023

All parts of this homework should be submitted to Canvas. Submit the answers to Question 1 and Question 2  using handwritten scanned notes or as a latex/typed up pdf file. Submit the answers  to Questions 3 and 4 as the pdf file generated when knitting this report. If need be, combine these pdfs into a single file

##Question 1: Comparing across eras
Assume the number of points per game each player scores in the nba is normally distributed with a different mean and standard deviation each season, that is, there are higher scoring eras and lower scoring eras.

World B. Free played in the 1970's and 1980's and averaged 24.1 points per game in the 1981 season. 
Metta World Peace played in the 2000's and 2010's and averaged 20.5 points per game in the 2008 season. 

(Both of these players were not born with these names, but legally changed their names to have a similar message.) 

Your grandfather tells you he watched both payers and he thought World B. Free in 1981 was much better than Metta World Peace in 2008. You suspect your grandfather is just a hater who thinks everything was better back in his day (he said Olivia Rodrigo is a poor man's Lorde playing Avril Lavigne B-sides, and he was TikTok famous back when it was musical.ly, and Vine was way better anyway.)     

You want to see if your grandpa has a point, so you find the distribution of points per game in 1981 was a Normal with a mean of 19 and a standard deviation of 3. The distribution of points per game in 2008 was a Normal with a mean of 16 and a standard deviation of 2.5. If all you care about is points, can you argue Metta World Peace was better? Why or why not?

We will use the z scores of the players to see which player is relatively farther from their respective means by standard deviation. We can do this because both distributions we are pulling data from are normal.

World B. Free's z-score = z1
Metta World Peace's z-score = z2

z = (x-mu)/sigma

z1 = (24.1-19)/3 = 1.7

z2 = (20.5-16)/2.5 = 1.8

In terms of points, Metta World Peace in 2008 was more superior than his peers than World B. Free in 1981 was.

##Question 2: Using the symmetry of the normal distribution
Without using R, assume the 68 - 95 - 99.7 rule holds precisely (not approximately).



###A)
What proportion of a Normal distribution falls within at least 2 standard deviations of the mean?

95% of the data.


###B) 
Using the symmetry of the Normal distribution, what percentage of the  Normal distribution falls below 1 standard deviation below the mean? What percentage of the data falls above 1 standard deviation above the mean? Explain your reasoning

Because 68% of the data is contained within 1 standard deviation below and above the mean, then 32% of this data is outside the first standard deviations. 
For each side: 32% / 2 = 16%
That means there is 16% of the data below 1 standard deviation below the mean and 16% of the data above 1 standard deviation above the mean.

###C) 
what percentage of the  Normal distribution falls either below 2 standard deviations below the mean or above 1 standard deviation above the mean? Explain your reasoning

Because 95% of the data falls within 2 standard deviations, 5% of the data fall below and above. So for one side: 5% / 2 = 2.5%

In part B, it was shown the 32% of the data falls above or below 1 standard deviation of the mean. So for one side: 32% / 2 = 16%

So, 2.5% + 17% = 18.5%

## Question 3: CDF values of Continuous RVs
###A
Use a geometric argument to calculate the probability a Uniform(0, 1) RV takes on a value less than .3

Because a Uniform(0,1) creates a rectangle from 0 to 1 with a uniform height of 1, then the area of the rectangle determines the the probability of a given value. 
So, the probability of a value falling between 0 and 0.3 is 0.3 or 30%

###B
Use R to find the probability a Uniform(0, 1) RV takes on a value less than .3
```{r}
print(punif(0.3, 0, 1))
```

###C
Use R to find the probability an Exponential(2) RV takes on a value less than .3
```{r}
print(pexp(0.3, 2))
```


###D
Use R to find the probability an Exponential(2) RV takes on a value greater than .5
```{r}
print(1 - pexp(0.5, 2))
```


###E
Use R to find the probability an Exponential(.5) RV takes on a value between .1 and .3
```{r}
print(pexp(0.3, 0.5) - pexp(0.1, 0.5))
```


###F
Use R to find the probability an Exponential(3) RV takes on a value either less than .4 or greater than 1
```{r}
print(pexp(0.4, 3) + (1 - pexp(1,3)))

```




##Question 4: Theoretical sample vs experimental sample 
We are checking whether a drug is effective in producing more of a particular protein. We administer the drug to 50 people. We know that in the general population, the parts per million of this particular protein in people's blood is normally distributed with a mean of 10 and a standard deviation of 5. After a month on the drug, the 50 people's blood was tested and the sample mean of the group was 10.3. 

###A
How likely is it that if we took a sample of 50 people from the general population, we would get a sample mean greater than 10.3? Calculate this theoretical value 

```{r}

SE <- 5 / sqrt(50)
z <- (10.3 - 10) / SE
print(1 - pnorm(z))


```

###B
Now do this empirically. Take 5000 samples of size 50 and tell me the probability that the sample mean exceeds 10.3

```{r}
num <- 5000
count <- 0 

for (i in 1:num) {
  mean <- mean(rnorm(50, 10, 5))
  if (mean > 10.3) {
    count <- count + 1
  }
}

print(count / num)

```


###C
What if we took a much larger sample. If instead we tested the drug on 10,000 people and got a sample mean of 10.25.  

How likely is it that if we took a similar sample from the general population, we would get a sample mean greater than 10.25? Calculate this value theoretically

```{r}
p_mean <- 10
p_sd <- 5
s_mean <- 10.25
s_size <- 10000

z <- (s_mean - p_mean) / (p_sd / sqrt(s_size))

print(1 - pnorm(z))
```

##D
Comment on the difference between the case with 50 people in the trial versus with 10,000 people in the trial. What practical difference is there? Would either of these results convince you the drug works? Did anything surprise you? Can you explain this result? 
When the trial was done with 10,000 people, the sample size is much larger and therefore it represents a more accurate estimation of the population mean. The result with 10,000 people would be much more convincing due to the fact that the z-score was very large. In the large sample, the probability of the sample mean being greater than 10.25 was nearly 0 when there was about a 0.33 chance of it being greater than 10.3 in the small sample trial. It is surprising to see how a larger sample size can essentially define the range of possibilities for a sample mean.
